---
title: "Injury imputation"
output: rmarkdown::github_document
---
  

```{r setup, include=FALSE}
library(knitr)
library(summarytools)
knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, message = F, warning = F, results='asis')

st_options(bootstrap.css     = FALSE,       # Already part of the theme so no need for it
           plain.ascii       = FALSE,       # One of the essential settings
           style             = "rmarkdown", # Idem.
           dfSummary.silent  = TRUE,        # Suppresses messages about temporary files
           footnote          = NA,          # Keeping the results minimalistic
           subtitle.emphasis = FALSE)       # For the vignette theme, this gives

load('outputs/workspace.Rdata')
#load('outputs/variables.Rdata')
#city_distances <- readRDS('processed_data/distances/city_distance.Rds')
#city_injuries <- readRDS('processed_data/whw_matrices/city_injuries.Rds')
```

```{r identify cities with and without whw matrices,echo=F}

## separate into those cities with whw and those without
cities_with_whw <- sapply(cities,function(x)'whw'%in%names(city_injuries[[x]]))
```


Cities `r paste0(cities[cities_with_whw],collapse=', ')` have who-hit-whom (WHW) matrices.

Cities `r paste0(cities[!cities_with_whw],collapse=', ')` don't.

Can we use the WHW matrices of the cities that have them and predict them for those that don't, using distances by mode for all cities?

The data are distances and fatal road-traffic collisions taken from http://github.com/ithim/ithim-r.


## Basic regression
We start with cities with WHW matrices, and we keep only injuries that have distances for both parties involved in the crash. 

```{r read table,echo=F}
injury_table <- readRDS('outputs/whw_table.Rds')
```

We fit a model with three categorical covariates: the mode of the casualty, the mode of the colliding vehicle, and the city. We assume a Poisson distribution for the fatality counts and use the distances travelled and the years of data collection as the offset.

The figure shows how much each level of the three covariates is expected to contribute to risk:

```{r plot,echo=F}
mod <- glm(count ~ mode1+mode2+city,data=injury_table,family=poisson(link=log),offset=0.5*log(dist1)+0.5*log(dist2)+log(years))
coefficients <- coef(mod)
nm <- names(coefficients)


identifiers <- c('mode1','mode2','city')
nms <- list()
vals <- list()
for(i in 1:length(identifiers)){
  mode1s <- grepl(identifiers[i],nm)
  m1_nms <- gsub(identifiers[i],'',nm[mode1s])
  nms[[i]] <- c(m1_nms,setdiff(unique(injury_table[[identifiers[i]]]),m1_nms))
  vals[[i]] <- c(coefficients[mode1s],0)
}

cols <- c('hotpink','grey','darkorange2')
par(mar=c(5,5,2,2))
plot(0,0,col='white',xlim=c(2,10.5),ylim=range(c(coefficients[-1])),frame=F,xaxt='n',xlab='',ylab='Danger',cex.axis=1.5,cex.lab=1.5)
for(i in 1:length(identifiers)){
  xval <- rep(3*i,length(nms[[i]]))
  yval <- vals[[i]]
  for(j in 2:length(yval))
    for(k in 1:(j-1)){
      if(abs(diff(yval[c(k,j)]))<0.2){
        xval[j] <- xval[j] + 0.5
        xval[k] <- xval[k] - 0.5
      }
    }
  text(labels=nms[[i]][yval!=0],x=xval[yval!=0],y=yval[yval!=0],col='navyblue')
  text(labels=nms[[i]][yval==0],x=xval[yval==0],y=yval[yval==0],col=cols[i])
}
axis(side=1,at=c(3,6,9),labels=c('Casualties','Colliders','Cities'),cex.axis=1.5)
```

The coloured levels in the middle are the reference values for each category. As we might expect: pedestrians are most at risk, and pose the least risk; trucks suffer the least risk. NOV, which stands for no other vehicle, seems to cause most fatalities - perhaps this represents missing information, rather than true collisions with no other vehicle.

*******************

## Evidence synthesis

To include the cities without matrices in the regression model, and to impute their matrices, we use the information we have: the number of fatalities per mode, each of which (in reality) is a sum over collider modes. This constitutes a separate arm of the inference, which again uses a Poisson distribution to describe the sums. The new cities' counts will drive the coefficients for the city intercepts; the regression coefficients for the collider modes from the old cities will drive the disaggregation of the sum into component collider parts; and all cities, new and old, will inform the coefficients for the casualty modes.

```{r compile pred table,echo=F}
coefficients <- coef(glm1)
nm <- names(coefficients)
coefficients <- results_list[,1] # coef(glm1)
names(coefficients) <- nm

identifiers <- c('mode1','mode2','city')
nms <- list()
vals <- list()
for(i in 1:length(identifiers)){
  mode1s <- grepl(identifiers[i],nm)
  m1_nms <- gsub(identifiers[i],'',nm[mode1s])
  nms[[i]] <- c(m1_nms,setdiff(unique(injury_table[[identifiers[i]]]),m1_nms))
  vals[[i]] <- c(coefficients[mode1s],0)
}

cols <- c('hotpink','grey','darkorange2')
par(mar=c(5,5,2,2))
plot(0,0,col='white',xlim=c(2,10.5),ylim=range(c(coefficients[-1])),frame=F,xaxt='n',xlab='',ylab='Danger',cex.axis=1.5,cex.lab=1.5)
for(i in 1:length(identifiers)){
  xval <- rep(3*i,length(nms[[i]]))
  yval <- vals[[i]]
  for(j in 2:length(yval))
    for(k in 1:(j-1)){
      if(abs(diff(yval[c(k,j)]))<0.2){
        xval[j] <- xval[j] + 0.5
        xval[k] <- xval[k] - 0.5
      }
    }
  text(labels=nms[[i]][yval!=0],x=xval[yval!=0],y=yval[yval!=0],col='navyblue')
  text(labels=nms[[i]][yval==0],x=xval[yval==0],y=yval[yval==0],col=cols[i])
}
axis(side=1,at=c(3,6,9),labels=c('Casualties','Colliders','Cities'),cex.axis=1.5)
```

We see a similar pattern as before in terms of the modes (indeed the collider coefficients are (and should be) exactly the same). The two new cities share the middle of space of all the cities with Belo Horizonte, Bangalore and Bogota.

***************************

## Imputed matrices

From the output of the regression, we can estimate the expected fatalities for each combination of modes:

### Buenos Aires
```{r predict ba,echo=F}
kable(round(ba_pred))
```


### Mexico City

```{r predict mc,echo=F}
kable(round(mc_pred))
```

*******************

### Who-hit-whom matrices

These can be compared to the raw matrices from the other cities:

```{r print whw,echo=F}
for(city in cities[cities_with_whw]) {
  cat(city)
  rounded <- round(city_injuries[[city]]$whw/city_weights[[city]])
  print(kable(rounded,caption=city))
    cat("\n")
}
```

********************

Still to do:

Include interactions between covariates (particularly pedestrian-pedestrian collisions); include distributions to describe distances; ascribe noise differentially to modes; 